Developing a public preprint culture in biology

### 2012-10-11 ms.tex
A bit about early career, and a few comments early in the text. Also minor edits, and a modification to the title

\title{Developing a preprint culture in biology}

@@

@@

Finally, public preprint servers offer a fair way to establish intellectual
priority by making the work available when it is complete.  Some manuscripts
will spend much more time in the review process and/or in production after
acceptance, than others. This means that publication and acceptance dates do not
accurately characterize who came up with an idea first.  For this reason,
mathematicians and physicists have embraced arXiv in part to establish priority
in a fair way \cite{gin11,cal12}.

@@

In contrast to other disciplines, the field of biology has effectively no
preprint culture, with the exception of small pockets of primarily highly
quantitative research (e.g., epidemiology). Submitting to preprint servers has
become more common in the past few years, and the quantitative biology section
in arXiv is experiencing the fastest growth in submissions \cite{cal12}. Yet,
the number of biology papers submitted to preprint servers still represents only
a small fraction of the total research produced in biology.

There are a number of reasons that biologists have not traditionally had a
culture of sharing preprints, many of which are based on common misconceptions
regarding the costs and benefits of sharing preprints. For example, in contrast
to other fields there is a perception in biology that public preprints
make it easier to steal ideas. However, there is no evidence of this happening
in the numerous other fields that have adopted preprint servers, and since preprint
servers create a clear record of who had the idea first, and when, this
appears to be a largely unfounded concern. In other fields preprints serve the
opposite role, they allow straightforward establishment of precedence, letting
research lay claim to an idea thus preventing it from being "stolen".

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice. This is
based on a certain interpretation of the Ingelfinger rule: scientists should not
publish the same manuscript twice \cite{alt96}. A preprint is simply a document
that allows ideas to spread and be discussed, it is not yet formally validated
by the peer-review system. This is why the majority of publishers do not see
arXiv and similar services as a violation of the Ingelfinger rule. Almost all of
the major publishers in biology are preprint-friendly, including: Nature
Publishing Group, PLOS, BMC, PNAS, Elsevier, and Springer \ref{table:policies}.
The Ecological Society of America and the Genetics Society of America also
recently changed their policies to allow public preprints. In fact,
\emph{Nature} responded to the rumour that they refused manuscripts submitted to
arXiv by saying that ``\emph{Nature} never wishes to stand in the way of
communication between researchers. We seek rather to add value for authors and
the community at large in our peer review, selection and editing'' \cite{nat05}.
Still, a few journals adopt a ``by default'' hostile attitude towards preprints,
mostly due to the lack of clear policy of the publishers, or perhaps because a
preprint culture has not developed in biology and the practice is still
considered unusual. As an example, Wiley-Blackwell, which publishes some of the
leading journals in biology, has no official policy on the matter.

>>>>

\title{Developing a public preprint culture in biology}

@@

%%% TP Oct 11 2012
%%% Just something I am thinking about now
%%% We are not a good sample of biologists. We are mostly evolutionary ecologists,
%%% and perhaps having the input of a "real" (labcoat wearing) biologist can help.

@@

Finally, public preprint servers offer a fair way to establish intellectual
priority by making the work available as soon as it is complete. Some
manuscripts will spend much more time in the review process and/or in
production after acceptance, than others. This means that publication and
acceptance dates do not accurately characterize who came up with an idea
first. For this reason, mathematicians and physicists have embraced arXiv in
part to establish priority in a fair way \cite{gin11,cal12}. This can be
especially crucial to early-career scientists. The sometimes long delay before
publication is seldom compatible with the pressure to show an impressive
publication record when applying for a scholarship or a position. Increasing
the perceived value of pre-prints as close, or equal, to journal articles (a
point we discuss in the conclusion) will allow young researchers to put their
research outcome in the open, and build a reputation for themselves through
the diffusion of their work, without the fear that this work will not be
recognized by grant or job commitees.

%%% TP Oct 11 2012
%%% Important to appeal to the young folks, while at the same time calling for
%%% more recognition of preprints. Feel free to rephrase, but don't remove!

@@

In contrast to other disciplines, the field of biology has effectively no
%%% TP
%%% As good a place as any to define biology... PLoS Biology, for example,
%%% is mostly cell biol stuff. We are not a typical subset of biologists!
preprint culture, with the exception of small pockets of primarily highly
quantitative research (\emph{e.g.}, epidemiology, population genetics).
Submitting to preprint servers has become more common in the past few years,
and the quantitative biology section in arXiv is experiencing the fastest
growth in submissions \cite{cal12}. Yet, the number of biology papers
submitted to preprint servers still represents only a small fraction of the
total research produced in biology.

There are a number of reasons that biologists have not traditionally had a
culture of sharing preprints, many of which are based on common misconceptions
regarding the costs and benefits associated with the practice. For example, in contrast
to other fields there is a perception in biology that public preprints
make it easier to steal ideas. However, there is no evidence of this happening
in the numerous other fields that have adopted preprint servers, and since preprint
servers create a clear record of who had the idea first, and when, this
appears to be a largely unfounded concern. In other fields preprints serve the
opposite role, they allow straightforward establishment of precedence, letting
research lay claim to an idea thus preventing it from being "stolen".
%%% TP
%%% Most refs will ask for a ref at that point. At least, I will, but I'm so
%%% demanding that I'm part of why the review process sucks ;-)

Another major concern is that if a researcher posts a public preprint that
they will not be allowed to submit their paper to their journal of choice.
This is based on a certain interpretation of the Ingelfinger rule: scientists
should not publish the same manuscript twice \cite{alt96}. A preprint is
simply a document that allows ideas to spread and be discussed, it is not yet
formally validated by the peer-review system. This is why the majority of
publishers do not see arXiv and similar services as a violation of the
Ingelfinger rule, and it is not unseen to have a preprint reproduced almost
exactly in the published version.
%%% TP
%%% Steve Frank papers on ArXiV often end up almost unmodified in JEB
%%% I can dig up the two version so that we can cite them
Almost all of the major publishers in
biology are preprint-friendly, including: Nature Publishing Group, PLOS, BMC,
PNAS, Elsevier, and Springer \ref{table:policies}. The Ecological Society of
America and the Genetics Society of America recently changed their
policies to allow public preprints. In fact,
\emph{Nature} responded to the rumour that they refused manuscripts submitted to
arXiv by saying that ``\emph{Nature} never wishes to stand in the way of
communication between researchers. We seek rather to add value for authors and
the community at large in our peer review, selection and editing'' \cite{nat05}.
Still, a few journals adopt a ``by default'' hostile attitude towards preprints,
mostly due to the lack of clear policy of the publishers, or perhaps because a
preprint culture has not developed in biology and the practice is still
considered unusual. As an example, Wiley-Blackwell, which publishes some of the
leading journals in biology, has no official policy on the matter.
